Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-01-03 WSSSPE4_report/wg_soft_eng.tex
addressed Dan's comments

\note{Anshu to write this}

%Introduction to group here, including the overall objective of work in this area.
This working group concerned itself with identifying processes that are not adequately
covered by general software engineering. The group focused
on testing because there is a growing awareness about the lack of trust
and reproducibility in many computational results. A rigorous
verification process that also involves ongoing regular testing is
fundamental to addressing the issue of trust. Regular automated
testing is the first step towards software processes that can lead to
provenance and reproducibility, the hallmarks of quality science.

\katznote{this paragraph repeats some of the previous one - they probably should be merged}
Computational science and engineering applications have many moving
parts that need to interoperate with one another. The accuracy and
reliability of results produced by the scientific software depends not
only on the individual components behaving correctly, but also on the
validity of their interactions. Therefore, a rigorous verification
process and a robust testing regime are critical requirements for
scientific software.  There have been instances where inadequate
verification has resulted in publication of wrong results and later
retraction of publications. As scientific understanding grows,
the corresponding computational software models are refined, leading
to more complex codes. Increasing complexity makes them more prone to
defects, not only in individual code units, but also in interaction
among units. Therefore, a strong verification process combined with a
rigorous testing regime plays a critical role in the prevention of
generating incorrect scientific results. However, most science teams
struggle to find a good solution for themselves. Causes range from
lack of exposure to the practices, to distrust of adopting practices
because they do not meet the needs of the teams developing such
software.

\subsubsection{Participants}
\begin{itemize}
\item Mark Abraham
\item Anshu Dubey %- is a computer scientist at the Mathematics and
%  Computer Science Division at Argonne National Laboratory. She has
%  been deeply involved with the development of multiphysics scientific
%  software in multiple domains which has lead to an interest in, and
%  understanding of, software engineering concerns in scientific
%  computing. 
\item Hans Fangohr
\item Dominic Kempf
\item Eric Seidel
\end{itemize}

@@

%How could a reader get more information or get more involved?
Readers interested in getting more information should get in touch
with a member of the working group. The working group
has a channel in the WSSSPE Slack. The channel is called
{\it wg-testing-in-science}.
(See \S\ref{sec:slack} for instructions on how to join the Slack WSSSPE team.)
Additionally, a git
repository \katznote{add url} exists for contributing content and reference to, and
curation of the existing literature on this topic.

>>>>

%\note{Anshu to write this}

%Introduction to group here, including the overall objective of work in this area.
This working group concerned itself with identifying processes that are not adequately
covered by general software engineering. Verification and testing is
the first subtopic that we address.  
% there is a growing awareness about the lack of trust
% and reproducibility in many computational results. 
%\katznote{this paragraph repeats some of the previous one - they probably should be merged}
Computational science and engineering applications have many moving
parts that need to interoperate with one another. The accuracy and
reliability of results produced by the scientific software depends not
only on the individual components behaving correctly, but also on the
validity of their interactions. 
% Therefore, a rigorous verification
% process and a robust testing regime are critical requirements for
% scientific software.  There have been instances where inadequate
% verification has resulted in publication of wrong results and later
% retraction of publications.
As scientific understanding grows
the corresponding computational software models are refined, leading
to more complex codes. Increasing complexity makes them more prone to
defects, not only in individual code units, but also in interaction
among units. Therefore, a strong verification process combined with a
rigorous testing regime plays a critical role in the prevention of
generating incorrect scientific results. However, most science teams
struggle to find a good solution for themselves. Causes range from
lack of exposure to the practices, to distrust of adopting practices
because they do not meet the needs of the teams developing such
software.  The current focus of our effort is on testing because it is the first step
towards building a software processes that can lead to provenance and
reproducibility, the hallmarks of quality science.

\subsubsection{Participants}
\begin{itemize}
\item Mark Abraham
\item Anshu Dubey 
\item Hans Fangohr
\item Dominic Kempf
\item Eric Seidel
\end{itemize}

@@

%How could a reader get more information or get more involved?
Readers interested in getting more information should get in touch
with a member of the working group. The working group
has a channel in the WSSSPE Slack. The channel is called
{\it wg-testing-in-science}.
(See \S\ref{sec:slack} for instructions on how to join the Slack WSSSPE team.)
Additionally, a git
repository (\url{https://github.com/WSSSPE/WG-Best-Practices.git}) exists for contributing content and reference to, and
curation of the existing literature on this topic.
